---
title: "Color Codes for Blackstone"
output: html_document
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
    fig.width = 8,
    fig.height = 6,
    out.width = "100%"
    )
# Prevents sci notation and sets the output of decimals to 4 (0.0000):
options(scipen = 999, digits = 4,
        tibble.print_min = 5, tibble.print_max = 5)

# Load Libraries:
library(colorspace)
library(magrittr)
library(tidyverse)
library(blackstone) # blackstone R package

set.seed(424429) # for reproducibility


######## Data
items <- dplyr::tibble(
  pre_Organization = c(1, 2, 3, 4, 5, 4, 3, 2, 1),
  post_Organization = dplyr::if_else(pre_Organization < 5, pre_Organization + 1, pre_Organization),
  pre_Source = c(2, 2, 3, 5, 4, 3, 2, 1, 2),
  post_Source = dplyr::if_else(pre_Source < 4, pre_Source + 2, pre_Source),
  pre_Publish = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
  post_Publish = pre_Publish + 2,
  pre_Write = c(2, 2, 2, 3, 3, 3, 4, 4, 4),
  post_Write = pre_Write + 1,
  pre_Research = c(1, 1, 2, 2, 3, 3, 4, 4, 4),
  post_Research = pre_Research + 1
)

items_single <- dplyr::tibble(
  Organization = c(1, 2, 3, 4, 5, 4, 3, 2, 1),
  Source = c(2, 2, 3, 5, 4, 3, 2, 1, 2),
  Publish = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
  Write = c(2, 2, 2, 3, 3, 3, 4, 4, 4),
  Research = c(1, 1, 2, 2, 3, 3, 4, 4, 4),
)

# Set scale_labels for recodeCat function:
# scale_labels as a named character vector, items in correct order:
levels_min_ext_named <- c(
  "Minimal" = "1", "Slight" = "2", "Moderate" = "3",
  "Good" = "4", "Extensive" = "5"
)

# levels_min_ext as just the names from levels_min_ext_named:
levels_min_ext <- names(levels_min_ext_named) 
# Question labels as a named vector with the naming structure
# like this: c("new label" = "original variable name"):
question_labels <- c(
  "Publish a lot of high quality papers" = "Publish",
  "Write a lot of research papers" = "Write",
  "Research in a lab with faculty" = "Research",
  "Organization of a large research project" = "Organization",
  "Source work for a research paper" = "Source"
)

# Recode the numeric to factor variables using the levels from levels_min_ext and Select the factor variables:
cat_items <- blackstone::recodeCat(items, levels_min_ext_named) %>% dplyr::select(dplyr::where(is.factor))
cat_items_single <- blackstone::recodeCat(items_single, levels_min_ext_named) %>% dplyr::select(dplyr::where(is.factor))


######## Data: Strongly disagree to Strongly Agree
items_single_div <- dplyr::tibble(
  Organization = c(1, 2, 3, 4, 5, 6, 7, 7, 6, 5, 4, 3, 2, 1, 1),
  Source = c(2, 2, 3, 5, 4, 3, 2, 1, 2, 6, 6, 6, 7, 7,7),
  Publish = c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4 , 5, 6, 7, 6),
  Write = c(2, 2, 2, 3, 3, 3, 4, 4, 4, 5,6,7,1,2,3),
  Research = c(1, 1, 2, 2, 3, 3, 4, 4, 4, 5,6,7,1,2,3),
)

items_div <- dplyr::tibble(
  pre_Organization = items_single_div[["Organization"]],
  post_Organization = dplyr::if_else(pre_Organization < 7, pre_Organization + 1, pre_Organization),
  pre_Source = items_single_div[["Source"]],
  post_Source = dplyr::if_else(pre_Source < 5, pre_Source + 2, pre_Source),
  pre_Publish = items_single_div[["Publish"]],
  post_Publish = dplyr::if_else(pre_Publish < 6, pre_Publish + 2, pre_Publish),
  pre_Write = items_single_div[["Write"]],
  post_Write = dplyr::if_else(pre_Write < 7, pre_Write + 1, pre_Write),
  pre_Research = items_single_div[["Research"]],
  post_Research = dplyr::if_else(pre_Research < 7, pre_Research + 1, pre_Research) 
)

# Set scale_labels for recodeCat function:
# scale_labels as a named character vector, items in correct order:
levels_dis_agree_named <- c(
  "Strongly Disagree" = "1", "Somewhat Disagree" = "2", "Disagree" = "3",
  "Neither Agree or Disagree" = "4", "Agree" = "5", "Somewhat Agree" = "6", "Strongly Agree" = "7"
)

# levels_min_ext as just the names from levels_min_ext_named:
levels_dis_agree <- names(levels_dis_agree_named) 

# Recode the numeric to factor variables using the levels from levels_min_ext and Select the factor variables:
cat_items_div <- blackstone::recodeCat(items_div, levels_dis_agree_named) %>% dplyr::select(dplyr::where(is.factor))
cat_items_single_div <- blackstone::recodeCat(items_single_div, levels_dis_agree_named) %>% dplyr::select(dplyr::where(is.factor))

```

```{css, echo = FALSE}
body, ul, li{
  font-family: "Arial";
  font-size: 20px;
}

h1, h2, h3, h4, h5, h6 {
  font-family: "Arial";
}

h1 {
  background-color: #2C2C4F;
  color: white;
  font-size: 38px;
  font-weight: 500;
}

h2 {
  background-color: #dadada;
  color: #283251;
  font-size: 32px;
  font-weight: bold;
}

h3 {
  background-color: #dadada;
  color: #18323a;
  font-size: 28px;
  font-weight: bold;
}
    
h4 {
  color: #283251; 
  font-size: 26px;
}
  
h5 {
  color: #18323a; 
  font-size: 24px;
}

hr {
  height: 3px;
  background-color: #283251;
}

```

# Updated Color Codes and Guidance for Data Visualizations

Updated guidance for colors to be used in all visuals for Blackstone Research and Evaluation

There are three new color palettes, each to be used in specific circumstances: 

  - The Blue Sequential Color Palette
  - The Blue-Red Diverging Color Palette
  - The Qualitative Color Palette

The Blue Sequential Color Palette should be used with stacked bar charts that have a likert scale that has one clear direction like: `r levels_min_ext`

The Blue-Red Diverging Color Palette should be used with stacked bar charts that have a likert scale that is folded or runs from a negative to positive valence like this: `r levels_dis_agree` 

The Qualitative Color Palette should be used for arrow charts with a grouping variable or horizontal bar charts with demographics (this palette may have limited uses at this time).

Below are the full hex color codes for each of the color palettes and examples that use each of color palettes.

Please reach out to Zack Crowley via Google Chat or email if you need hex codes for scales or different lengths than those provided here or 
if you have any other questions or clarifications about which color palettes to use and data visualizations in general. 

## Blue Sequential Color Palette

### Hex Codes for the Blue Sequential Color Palette

```{r, results='asis'}
cat("<h4><b>5 point scale:</b></h4>")
cat(paste0(paste0("`", seqFillColors(n_colors = 5), "`"), collapse = ", "))
cat("<br>")

```

```{r fig.width = 5,fig.height = 1}
colors <- seqFillColors(n_colors = 5)
labels_text_color <- labelColorMaker(colors = colors)
ggplot2::ggplot(data.frame(id = rev(seq_along(colors)), color = rev(colors))) + 
  ggplot2::geom_col(ggplot2::aes(y = 1, x = id, fill = rev(colors)), color = "black", width = 1) + 
  ggplot2::geom_text(ggplot2::aes(y = 1, x = id, label = rev(colors)), position = ggplot2::position_fill(vjust = 0.5), color = rev(labels_text_color)) + 
  ggplot2::scale_fill_identity() +
  ggplot2::scale_color_identity() +
  ggplot2::theme_void()
```

```{r, results='asis'}
cat("<br>")
cat("<h4><b>7 point scale:</b></h4> ")
cat(paste0(paste0("`", seqFillColors(n_colors = 7), "`"), collapse = ", "))
cat("<br>") 

```

```{r fig.width = 7,fig.height = 1.25}
colors <- seqFillColors(n_colors = 7)
labels_text_color <- labelColorMaker(colors = colors)
ggplot2::ggplot(data.frame(id = rev(seq_along(colors)), color = rev(colors))) + 
  ggplot2::geom_col(ggplot2::aes(y = 1, x = id, fill = rev(colors)), color = "black", width = 1) + 
  ggplot2::geom_text(ggplot2::aes(y = 1, x = id, label = rev(colors)), position = ggplot2::position_fill(vjust = 0.5), color = rev(labels_text_color)) + 
  ggplot2::scale_fill_identity() +
  ggplot2::scale_color_identity() +
  ggplot2::theme_void()
```

### Example use of the Blue Sequential Color Palette in a Stacked Bar Chart with a 5-point Likert Scale:

```{r}
blackstone::stackedBarChart(
   df = cat_items_single, pre_post = FALSE, scale_labels = levels_min_ext,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)

```



<hr>


## The Blue-Red Diverging Color Palette

### Hex Codes for the Blue-Red Diverging Color Palette

```{r, results='asis'}
cat("<h4><b>5 point scale:</b></h4>")
cat(paste0(paste0("`", divFillColors(n_colors = 5), "`"), collapse = ", "))
cat("<br>")

```

```{r fig.width = 5,fig.height = 1}
colors <- divFillColors(n_colors = 5)
labels_text_color <- labelColorMaker(colors = colors)
ggplot2::ggplot(data.frame(id = rev(seq_along(colors)), color = rev(colors))) + 
  ggplot2::geom_col(ggplot2::aes(y = 1, x = id, fill = rev(colors)), color = "black", width = 1) + 
  ggplot2::geom_text(ggplot2::aes(y = 1, x = id, label = rev(colors)), position = ggplot2::position_fill(vjust = 0.5), color = rev(labels_text_color)) + 
  ggplot2::scale_fill_identity() +
  ggplot2::scale_color_identity() +
  ggplot2::theme_void()
```

```{r, results='asis'}
cat("<br>")
cat("<h4><b>7 point scale:</b></h4> ")
cat(paste0(paste0("`", divFillColors(n_colors = 7), "`"), collapse = ", "))
cat("<br>") 

```

```{r fig.width = 7, fig.height = 1.25}
colors <- divFillColors(n_colors = 7)
labels_text_color <- labelColorMaker(colors = colors)
ggplot2::ggplot(data.frame(id = rev(seq_along(colors)), color = rev(colors))) + 
  ggplot2::geom_col(ggplot2::aes(y = 1, x = id, fill = rev(colors)), color = "black", width = 1) + 
  ggplot2::geom_text(ggplot2::aes(y = 1, x = id, label = rev(colors)), position = ggplot2::position_fill(vjust = 0.5), color = rev(labels_text_color)) + 
  ggplot2::scale_fill_identity() +
  ggplot2::scale_color_identity() +
  ggplot2::theme_void()
```

### Example of a Stacked Pre-Post bar chart with The Blue-Red Diverging Color Palette on a 7 point scale:

```{r}
cat_items_div %>%
blackstone::stackedBarChart( 
   df = ., fill_color = "div", pre_post = T, scale_labels = levels_dis_agree,
   question_labels = question_labels, question_order = F, percent_label = TRUE, width = NULL, overall_n = T
)
```


<hr>


### Main color for arrow charts without groups or table headings

### Blackstone Dark Blue Hex Code `r blackstone::blackstoneColors["dark_blue"]`

```{r, fig.height = 2, fig.width = 1.5, out.width="20%", fig.align='center'}
# Create a data frame with the colors and their labels
blue_color_data <- tibble(
  color = blackstone::blackstoneColors["dark_blue"],
  label = "blackstone\n dark blue",
  labels_text_color = labelColorMaker(colors = blackstone::blackstoneColors["dark_blue"])
)
# Set up ggplot and resize using the R code chunk: {fig.width = 11, fig.height = 2}
ggplot2::ggplot(blue_color_data, aes(x = label, y = 1, fill = color)) +
  ggplot2::geom_col(color = "black", width = 1) +
  ggplot2::geom_text(aes(label = color), position = ggplot2::position_fill(vjust = 0.5), color = blue_color_data[["labels_text_color"]]) +
  ggplot2::scale_fill_identity() +
  ggplot2::theme_void() +
  ggplot2::theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

```


### Arrow chart example without grouping variable:

```{r}
blackstone::arrowChart(df = items, scale_labels = levels_min_ext,
           overall_n = T, question_labels = question_labels, question_order = TRUE)
```


## Qualitative/Categorical Color Palette

The Qualitative Color Palette should be used for arrow charts with a grouping variable or for horizontal bar charts for demographics.

### Hex codes for the Qualitative Color Palette

```{r, results='asis'}
# knitr::combine_words(customCols(), before = '"', after = '"', and = "", sep = ", ")
cat(paste0(paste0("`", qualColors(), "`"), collapse = ", "))
```


### Example Swatch of Qualitative Color Palette with names and hex codes:

```{r fig.width = 11, fig.height = 2}
# purrr::map_chr(seq_along(qualColors(add_names = T)), \(x) paste0("'",names(qualColors(add_names = T))[x], "':  ", qualColors(add_names = T)[x]))

# function to show color, names and hex codes as visual:
# show_colors2 <- function(colors) { 
#   labels_color <- purrr::map_chr(seq_along(colors), \(x) paste0(" '",names(colors)[x], "':\n", colors[x]))
#   labels_text_color <- labelColorMaker(colors = colors)
#   ggplot2::ggplot(data.frame(id = rev(seq_along(colors)), color = rev(colors))) + 
#     ggplot2::geom_tile(ggplot2::aes(1, id, fill = rev(color))) + 
#     ggplot2::geom_text(ggplot2::aes(1, id, label = labels_color), color = labels_text_color) + 
#     ggplot2::scale_fill_identity() +
#     ggplot2::scale_color_identity() +
#     ggplot2::theme_void()
# }

# Create a data frame with the colors and their labels
color_data <- tibble(
  color = qualColors(add_names = T),
  label = factor(names(qualColors(add_names = T)), levels = names(qualColors(add_names = T))),
  labels_text_color = labelColorMaker(colors = qualColors(add_names = T))
)
# Set up ggplot and resize using the R code chunk: {fig.width = 11, fig.height = 2}
ggplot2::ggplot(color_data, aes(x = label, y = 1, fill = color)) +
  ggplot2::geom_col(color = "black", width = 1) +
  ggplot2::geom_text(aes(label = color), position = ggplot2::position_fill(vjust = 0.5), color = color_data[["labels_text_color"]]) +
  ggplot2::scale_fill_identity() +
  ggplot2::theme_void() +
  ggplot2::theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```



### Example of an Arrow chart with Grouping Variable and the Qualitative Color Palette:

#### - The "Overall" group will always be the color black
#### - This also shows what left/negative direction arrows would look like with the `Research` items.

```{r}
# add a grouping var to the data set `items`: 
items_group <- dplyr::tibble(
   items,
   edu_level = factor(c(
       "grad", "undergrad", "grad", "undergrad", "grad",
       "undergrad", "undergrad", "grad", "undergrad"
  ), levels = c("grad", "undergrad"))
 ) %>% mutate(
            post_Research = dplyr::if_else(post_Research > 2, pre_Research - 2, pre_Research)
 ) 

blackstone::arrowChartGroup(df = items_group, group = "edu_level", group_levels = c("grad", "undergrad"), scale_labels = levels_min_ext, 
    overall_n = T, question_labels = NULL, question_order = FALSE)

```

### Example of a Horizontal Bar Chart with Qualitative Color Palette:

```{r}
# New fake data for testing horzBarChart() colors with racial categories:
# Generate fake data
n <- 100  # number of observations

# Define racial categories
racial_categories <- c("Asian", "Hispanic or Latino", "Caucasian or White (non-Hispanic)", "African American or Black", "Native American or American Indian", 
                       "Native Hawaiian or other Pacific Islander", "Mixed Race")

# Set up the fake racial data as a tibble:
data <- dplyr::tibble(
  race = sample(racial_categories, n, replace = TRUE) # Generate random racial categories for each observation
)

# set up data for chart:
race_summ <- data %>%
  dplyr::select(race) %>%
  blackstone::dataSumm(sort_n = T)

# Filtered to only three categories:
three_race_summ <- race_summ %>% filter(response %in% c("Asian", "Hispanic or Latino", "African American or Black")) 

r_colors <- blackstone::qualFillColors(n_colors = length(unique(race_summ[["response"]]))) %>% purrr::set_names(race_summ[["response"]])
# horzBarChart() with custom qual colors c("#E69F00"", "#56B4E9", "#009E73", "#CC79A7", "#D55E00", "#0072B2", "#440154FF")
blackstone::horzBarChart(df = race_summ, scale_colors = r_colors, width = 0.6)

```

<hr>



