---
title: "Importing and Cleaning Data"
author: "Blackstone Research and Evaluation"
date: "`r Sys.Date()`"
output: html_document
---

```{r here_i_am}
# Set here() location:
here::i_am("report.Rmd")
```

```{r setup, include = FALSE}
knitr::opts_chunk[["set"]]( 
    collapse = TRUE,
    comment = "#>",
    warning = FALSE, 
    message = FALSE, 
    fig.width = 9,
    fig.height = 8,
    out.width = "100%"
    )
# Prevents sci notation and sets the output of decimals to 4 (0.0000):
options(scipen = 999, digits = 4,
        tibble.print_min = 5, tibble.print_max = 5)

# Set seed to reproduce data:
set.seed(424629)
```

```{r load_libraries}
library(dplyr) # data manipulation
library(here) # file path management
library(blackstone) # Blackstone helper functions
```

```{r set_paths}
# Set up file paths:
# Helpful function to show folder structure:
# fs::dir_tree(here::here("..","..","..")) # shows the folder structure of `2429_DEMO/Phase2_Evaluation` folder
data_fp <- here::here("..","..","..", "Data") # Set up file path to 'Data" folder:
data_yr4_fp <- here::here(data_fp, "Year4_(2023-2024)")
list.files(data_yr4_fp, recursive = T)
```

## Pre Survey Data

```{r create_codebook}
# File path for the year 4 pre data:
pre_data_fp <- here(data_yr4_fp, "pre_data/sm_data_pre.csv")
# 1. Create the codebook:
codebook_pre <- blackstone::createCodebook(pre_data_fp)
codebook_pre

# Step 2. Edit the codebook: 
# Set up sequential naming convections for matrix-style questions with shared likert scale response options:
# 8 items that are matrix-style likert scales- turned into a scale called `research`- here is how to easily name them all at once:
# Rows 11 to 18 belong to the "research" matrix question (you will have to look at the codebook and match the header_1 and header_2 to variable_name to change)
research_items <- codebook_pre[["variable_name"]][11:18]
research_names <- paste0("research_", seq_along(research_items)) %>% purrr::set_names(., research_items) # Create a new named vector of names for these columns
# 6 items that are matrix-style likert scales- turned into a scale called `ability`- Rows 19 to 24 named `variable_name`:
ability_items <- codebook_pre[["variable_name"]][19:24]
ability_names <- paste0("ability_", seq_along(ability_items)) %>% purrr::set_names(., ability_items) # Create a new named vector of names for these columns
# 6 items that are matrix-style likert scales- turned into a scale called `ethics`- Rows 19 to 24 named `variable_name`:
ethics_items <- codebook_pre[["variable_name"]][25:29]
ethics_names <- paste0("ethics_", seq_along(ethics_items)) %>% purrr::set_names(., ethics_items) # Create a new named vector of names for these columns
# Edit the `variable_names` column: Use dplyr::mutate() and dplyr::case_match() to change the column `variable_name`:
codebook_pre <- codebook_pre %>% dplyr::mutate(
    variable_name = dplyr::case_match(
        variable_name, # column to match
        'custom_data_1' ~ "unique_id", # changes 'custom_data_1' to "unique_id"
        'to_what_extent_are_you_knowledgeable_in_conducting_research_in_your_field_of_study' ~ "knowledge",
        'with_which_gender_do_you_most_closely_identify' ~ "gender",
        'which_race_ethnicity_best_describes_you_please_choose_only_one' ~ "ethnicity",
        'are_you_a_first_generation_college_student' ~ "first_gen",
        names(research_names) ~ research_names[variable_name], # takes the above named vector and when the name matches, applies new value in that position as replacement.
        names(ability_names) ~ ability_names[variable_name],   # Same for `ability_names`
        names(ethics_names) ~ ethics_names[variable_name],   # Same for `ability_names`
        .default = variable_name # returns default value from original `variable_name` if not changed.
        )
    )
# codebook_pre

# Write out the edited codebook to save for future use-
# Be sure to double check questions match new names before writing out:
# readr::write_csv(codebook_pre, file = "{filepath-to-codebok}")

# 3. Read in the data and rename the vars using readRenameData(), passing the file path and the edited codebook:
pre_data <- blackstone::readRenameData(pre_data_fp, codebook = codebook_pre)
pre_data
```


## Post Survey Data

- Do that same process over again with the post data, if the variables are all the same you can use the same codebook. 

- For this example there are additional post variables so a new codebook will need to be created to rename the variables when reading in the data with `readRenameData()`.

```{r import_post_data}
# File path for pre example data:
post_data_fp <- blackstone::blackstoneExample("sm_data_post.csv")
# 1. Create the codebook using the filepath:
codebook_post <- blackstone::createCodebook(post_data_fp)

# Step 2. Edit the codebook: 
# Set up sequential naming convections for matrix-style questions with shared likert scale response options:
# 8 items that are matrix-style likert scales- turned into a scale called `research`- here is how to easily name them all at once:
# Rows 11 to 18 belong to the "research" matrix question (you will have to look at the codebook and match the header_1 and header_2 to variable_name to change)
research_items <- codebook_post[["variable_name"]][11:18]
research_names <- paste0("research_", seq_along(research_items)) %>% purrr::set_names(., research_items) # Create a new named vector of names for these columns
# 6 items that are matrix-style likert scales- turned into a scale called `ability`- Rows 19 to 24 named `variable_name`:
ability_items <- codebook_post[["variable_name"]][19:24]
ability_names <- paste0("ability_", seq_along(ability_items)) %>% purrr::set_names(., ability_items) # Create a new named vector of names for these columns
# 6 items that are matrix-style likert scales- turned into a scale called `ethics`- Rows 19 to 24 named `variable_name`:
ethics_items <- codebook_post[["variable_name"]][25:29]
ethics_names <- paste0("ethics_", seq_along(ethics_items)) %>% purrr::set_names(., ethics_items) # Create a new named vector of names for these columns
# 5 items that are Open-ended follow up when corresponeding ethics items were answered "Strongly disagree"or "Disagree"- Rows 30 to 34 named `variable_name`:
ethics_items_oe <- codebook_post[["variable_name"]][30:34]
ethics_names_oe <- paste0("ethics_", seq_along(ethics_items), "_oe") %>% purrr::set_names(., ethics_items_oe) # Create a new named vector of names for these columns
# Edit the `variable_names` column: Use dplyr::mutate() and dplyr::case_match() to change the column `variable_name`:
codebook_post <- codebook_post %>% dplyr::mutate(
    variable_name = dplyr::case_match(
        variable_name, # column to match
        'custom_data_1' ~ "unique_id", # changes 'custom_data_1' to "unique_id"
        'to_what_extent_are_you_knowledgeable_in_conducting_research_in_your_field_of_study' ~ "knowledge",
        'with_which_gender_do_you_most_closely_identify' ~ "gender",
        'which_race_ethnicity_best_describes_you_please_choose_only_one' ~ "ethnicity",
        'are_you_a_first_generation_college_student' ~ "first_gen",
        names(research_names) ~ research_names[variable_name], # takes the above named vector and when the name matches, applies new value in that position as replacement.
        names(ability_names) ~ ability_names[variable_name],   # Same for `ability_names`
        names(ethics_names) ~ ethics_names[variable_name],   # Same for `ability_names`
        names(ethics_names_oe) ~ ethics_names_oe[variable_name],   # Same for `ethics_names_oe`
        .default = variable_name # returns default value from original `variable_name` if not changed.
        )
    )

# Write out the edited codebook to save for future use-
# Be sure to double check questions match new names before writing out:
# readr::write_csv(codebook_post, file = "{filepath-to-codebok}")
# 3. Read in the data and rename the vars using readRenameData(), passing the file path and the edited codebook:
post_data <- blackstone::readRenameData(post_data_fp, codebook = codebook_post)
post_data
```

Finally, it is important to add `pre_` and `post_` prefixes to all unique variables before merging the datasets, (i.e. the survey items that differ pre-post- the SM items and demos are identical):

```{r}
# Pre data:
pre_data <- pre_data %>% dplyr::rename_with(~ paste0("pre_", .), .cols = c(knowledge:ethics_5))
# Pre data:
post_data <- post_data %>% dplyr::rename_with(~ paste0("post_", .), .cols = c(knowledge:ethics_5_oe))
```


## Merging Data

```{r merge_data}
# left_join() will automatically join by all the shared columns, be sure to include all shared variables that should be identical pre-post to the 'by = join_by()' as an arg
# (otherwise you will get a message about additional variables to be joined by):
sm_data <- post_data %>% dplyr::left_join(pre_data, by = dplyr::join_by(respondent_id, collector_id, start_date, end_date, ip_address, email_address, 
                                                                        first_name, last_name, unique_id, gender, ethnicity, first_gen))
sm_data
```

## Write Out Clean Data

Create a new directory for clean data in the Year 4 folder, if it doesn't exist:

```{r clean_dir}
if (!fs::dir_exists(here(data_yr4_fp, "clean_data"))) {fs::dir_create(here(data_yr4_fp, "clean_data"))}
```

Write out clean data using `here()` to set up the file path and name the file to be written as "sm_data_clean.csv":

```{r write_clean}
readr::write_csv(sm_data, file = here(data_yr4_fp, "clean_data", "sm_data_clean.csv"))
```


