---
title: "Color Tests for BRE"
output: html_document
date: "2024-05-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
    fig.width = 8,
    fig.height = 6)

# Load Libraries:
library("colorspace")

# color setup for new BRE colors:
bre_blue <- "#283251"
bre_grey <- "#c0bfbf"
bre_light_grey <- "#eaeaeb"
bre_med_grey <- "#cecece"

bre_colors <- c("dark_blue" = "#283251",
                "light_grey" = "#eaeaeb",
                "med_grey" = "#cecece",
                "main_grey" = "#c0bfbf")

# Trying out a white to blue palette:
pal_bre_blues <- colorRampPalette(colors = c("white", "#283251"))
pal_bre_blues_twelve <- pal_bre_blues(12)
scales::show_col(pal_bre_blues(5))
scales::show_col(pal_bre_blues(6))
scales::show_col(pal_bre_blues(7))
scales::show_col(pal_bre_blues(20))
scales::show_col(pal_bre_blues_twelve[1:12])
scales::show_col(pal_bre_blues(7)[3:7])
```

### Possible color scheme reccomendations for new BRE R package and all visuals

#### Single Sequential colors:

- These would be for likert scale items that have a single order like "Minimal" to "Extensive" 

- `BRE Blue` is the blue color for our branding to white, dropping the lightest shades.

##### Seven point scales:

```{r}
# Create a single swatchplot to share with Eval Management Team
swatchplot(
    "BRE Blue" = pal_bre_blues(9)[3:9],
    "Darker Blues"  = darken(sequential_hcl(7, "Blues", rev = TRUE)),
    "Blues 3"  = sequential_hcl(7, "Blues 3", rev = TRUE),
    "Oslo"  = sequential_hcl(7, "oslo")
)

```

##### Five point scales:

```{r}
# Create a single swatchplot to share with Eval Management Team
swatchplot(
    "BRE Blue" = pal_bre_blues(9)[5:9],
    "Darker Blues"  = darken(sequential_hcl(5, "Blues", rev = TRUE)),
    "Blues 3"  = sequential_hcl(5, "Blues 3", rev = TRUE),
    "Oslo"  = sequential_hcl(5, "oslo")
)

```

#### Single Sequential (Multi-hue) colors:

- These would be for likert scale items that have a single order like "Minimal" to "Extensive", but have a bit more color to break up 

- The `Viridis` would also work for qualitative/categories

```{r}
# Create a single swatchplot to share with Eval Management Team
swatchplot(
    "YlGnBu (Yellow Green Blue)"  = sequential_hcl(7, "YlGnBu", rev = TRUE),
    "GnBu"  = sequential_hcl(7, "GnBu", rev = TRUE),
    "Viridis" = sequential_hcl(7, "Viridis", rev = TRUE),
)

```

#### Diverging colors:

- These would mostly be for likert scale items that are folded scales like 'Strongly Disagee' to 'Strongly Agree'.

- The `Cividis` palette could work for any type of likert scale items stacked bar charts.

```{r}
# Create a single swatchplot to share with Eval Management Team
swatchplot(
    "Lighter Cividis" = lighten(divergingx_hcl(7, "cividis", rev = TRUE), amount = 0.2, method = "relative", space = "HCL"),
    "Cividis" = divergingx_hcl(7, "cividis", rev = TRUE),    
    "Geyser" = divergingx_hcl(7, "geyser", rev = TRUE),
    "Darker Geyser" = darken(divergingx_hcl(7, "geyser", rev = TRUE), amount = 0.2, method = "relative", space = "HCL"),
    "Blue-Red 3"  = diverging_hcl(7, "Blue-Red 3", rev = TRUE)
)

```



