---
title: "Color Blind Palette Tests for BRE"
output: html_document
date: "2024-05-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
    fig.width = 8,
    fig.height = 6)

# Load Libraries:
library("colorspace")
library("magrittr")

# color setup for new BRE colors:
bre_blue <- "#283251"
bre_grey <- "#c0bfbf"
bre_light_grey <- "#eaeaeb"
bre_med_grey <- "#cecece"

bre_colors <- c("dark_blue" = "#283251",
                "light_grey" = "#eaeaeb",
                "med_grey" = "#cecece",
                "main_grey" = "#c0bfbf")

# Trying out a white to blue palette:
pal_bre_blues <- colorRampPalette(colors = c("white", "#283251"))
pal_bre_blues_twelve <- pal_bre_blues(12)
scales::show_col(pal_bre_blues(5))
scales::show_col(pal_bre_blues(6))
scales::show_col(pal_bre_blues(7))
scales::show_col(pal_bre_blues(20))
scales::show_col(pal_bre_blues_twelve[1:12])
scales::show_col(pal_bre_blues(7)[3:7])
```

### Possible color scheme reccomendations for new BRE R package and all visuals

### All Color-blind Friendly Palettes:

#### My recommendations:
- "Cividis" as the default sequential color scale of likert items like: Minimal, Slight, Moderate, Good, Extensive- that are not folded/go in a single direction.
- "Blue-Red 3" as the default sequential color scale of likert items that are folded/have a positive and negative connotation like: Strongly disagree to Strongly Agree
- "Viridis" or "Okabe-Ito" as the default qualitative/categorical scales (e.g. demographics, gender, role).


```{r, include=FALSE}
# palette.pals() # get default palette names
# Generate palettes for 3 to 7 scales:
values <- c(3:7)
colorblind_swatches <- purrr::map(.x = values,  ~ swatchplot(
                                                             "Blue-Red 3"  = colorspace::diverging_hcl(n = .x, "Blue-Red 3", rev = TRUE),
                                                             "Mako" = viridis::viridis(n = .x, alpha = 1, begin = 0, end = 1, direction = -1, option = "mako"),
                                                             "Cividis" = viridis::viridis(n = .x, alpha = 1, begin = 0, end = 1, direction = -1, option = "cividis"),
                                                             # "Cividis Lite" = viridisLite::cividis(n = .x, alpha = 1, begin = 0, end = 1, direction = -1),
                                                             "Viridis" = viridis::viridis(n = .x, alpha = 1, begin = 0, end = 1, direction = -1, option = "viridis"),
                                                             "Okabe-Ito" = rev(palette.colors(n = .x, palette = "Okabe-Ito", recycle = FALSE, names = TRUE)),
                                                             )
) %>% purrr::set_names(values)

```

<!-- Print out the swatches -->

```{r, results='asis'}
  cat("<br>")
  purrr::walk(values, \(x) {
      cat("<br>") # line break
      cat(stringr::str_glue("<h5>Colorblind Friendly Swatches for **{x}** Items</h5>")) # h4 header
      # print:
      swatchplot(
                 "Blue-Red 3"  = colorspace::diverging_hcl(n = x, "Blue-Red 3", rev = TRUE),
                 "Mako" = viridis::viridis(n = x, alpha = 1, begin = 0, end = 1, direction = -1, option = "mako"),
                 "Cividis" = viridis::viridis(n = x, alpha = 1, begin = 0, end = 1, direction = -1, option = "cividis"),
                 # "Cividis Lite" = viridisLite::cividis(n = x, alpha = 1, begin = 0, end = 1, direction = -1),
                 "Viridis" = viridis::viridis(n = x, alpha = 1, begin = 0, end = 1, direction = -1, option = "viridis"),
                 "Okabe-Ito" = rev(palette.colors(n = x, palette = "Okabe-Ito", recycle = FALSE, names = TRUE)),
                 )
      cat("<br>") # line break
      # HR line to separate item sections
      cat("<hr>")
    }
  )
```
