---
title: "Color Tests for BRE"
output: html_document
date: "2024-05-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
    fig.width = 8,
    fig.height = 6)

# Load Libraries:
library(colorspace)
library(magrittr)
library(tidyverse)

###### colors:
bre_colors <- c("dark_blue" = "#283251",
                "light_grey" = "#eaeaeb",
                "med_grey" = "#cecece",
                "main_grey" = "#c0bfbf")

c("#283251", "#405683", "#62759a", "#a1aec4", "#e9ebf0")

# light grey to BRE Blue color scale:
pal_bre_grey_blue <- colorRampPalette(c(bre_colors["light_grey"], bre_colors["dark_blue"]))
# pal_bre_grey_blue(length(scale_labels))
# 
# white to BRE Blue color scale:
pal_bre_blue <- colorRampPalette(c("white", bre_colors["dark_blue"]))
# pal_bre_grey_blue(length(scale_labels))

######### Data
items <- dplyr::tibble(
  pre_Organization = c(1, 2, 3, 4, 5, 4, 3, 2, 1),
  post_Organization = dplyr::if_else(pre_Organization < 5, pre_Organization + 1, pre_Organization),
  pre_Source = c(2, 2, 3, 5, 4, 3, 2, 1, 2),
  post_Source = dplyr::if_else(pre_Source < 4, pre_Source + 2, pre_Source),
  pre_Publish = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
  post_Publish = pre_Publish + 2,
  pre_Write = c(2, 2, 2, 3, 3, 3, 4, 4, 4),
  post_Write = pre_Write + 1,
  pre_Research = c(1, 1, 2, 2, 3, 3, 4, 4, 4),
  post_Research = pre_Research + 1
)

items_single <- dplyr::tibble(
  Organization = c(1, 2, 3, 4, 5, 4, 3, 2, 1),
  Source = c(2, 2, 3, 5, 4, 3, 2, 1, 2),
  Publish = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
  Write = c(2, 2, 2, 3, 3, 3, 4, 4, 4),
  Research = c(1, 1, 2, 2, 3, 3, 4, 4, 4),
)

# Set scale_labels for recodeCat function:
# scale_labels as a named character vector, items in correct order:
levels_min_ext_named <- c(
  "Minimal" = "1", "Slight" = "2", "Moderate" = "3",
  "Good" = "4", "Extensive" = "5"
)

# bar_scale_labels as just the names from levels_min_ext:
bar_scale_labels <- names(levels_min_ext_named)

levels_min_ext <- names(levels_min_ext_named) 
# Question labels as a named vector with the naming structure
# like this: c("new label" = "original variable name"):
question_labels <- c(
  "Publish a lot of high quality papers" = "Publish",
  "Write a lot of research papers" = "Write",
  "Research in a lab with faculty" = "Research",
  "Organization of a large research project" = "Organization",
  "Source work for a research paper" = "Source"
)

# Recode the numeric to factor variables using the levels from levels_min_ext:
cat_items <- bre::recodeCat(items, levels_min_ext_named)
cat_items_single <- bre::recodeCat(items_single, levels_min_ext_named)

# Select the factor variables:
cat_items <- cat_items %>% dplyr::select(dplyr::where(is.factor))
cat_items_single <- cat_items_single %>% dplyr::select(dplyr::where(is.factor))
```


```{r, include=FALSE}
material_BRE_blue <- rev(c("#283251", "#384c77", "#62759a", "#a1aec4", "#e9ebf0"))

# Used material design app to generate colors, based off of BRE Blue (#283251) as primary: https://m2.material.io/inline-tools/color/ 
# Generates PRIMARY, COMPLEMENTARY, 2 ANALOGOUS, and 2 TRIADIC Colors for the palette:
# Set names like in md app:
names_nums_pals <- c("900", "800", "700", "600", "500", "400", "300", "200", "100", "50")
# Primary: 900 is main (BRE Blue, "#283251")
md_bre_blue <- c("#283251", "#32436a", "#384c77", "#405683", "#475e8c", "#62759a", "#7e8eab", "#a1aec4", "#c5cedc", "#e9ebf0") %>% purrr::set_names(names_nums_pals)
# COMPLEMENTARY: 800 is main ("#514728")
md_bre_comp <- c("#2e2505", "#514728", "#726646", "#877a59", "#b2a481", "#d1c29e", "#f4e4bf", "#fff0cb", "#fff5cf", "#fffad4") %>% purrr::set_names(names_nums_pals)
# ANALOGOUS: 800 is main for teal ("#284751"), 900 is main for teal ("#322851")
md_bre_teal <- c("#18323a", "#284751", "#355a66", "#436e7b", "#4f7d8c", "#68909e", "#82a4b1", "#a3bfc9", "#c3dae0", "#e2f1f8") %>% purrr::set_names(names_nums_pals)
md_bre_purple <- c("#322851", "#403869", "#484175", "#514b81", "#57528a", "#6d6b9a", "#8586ab", "#a6a8c4", "#c9cadc", "#e9eaf0") %>% purrr::set_names(names_nums_pals)

# Triadic: 900 is main for both ("#514728") and ("#512832")
md_bre_tri_purp <- c("#472851", "#603166", "#6e3670", "#7d3d79", "#88427f", "#9a5791", "#ac72a4", "#c498be", "#dbc0d7", "#f0e6ee") %>% purrr::set_names(names_nums_pals)
md_bre_maroon <- c("#512832", "#63353e", "#734147", "#854c53", "#92555b", "#a56f73", "#b9898b", "#d2a9ac", "#ebcacb", "#ffe7e5") %>% purrr::set_names(names_nums_pals)


# pal_bre_blue_md_gray <- colorRampPalette(c("#e9ebf0", bre_colors["dark_blue"]))

swatchplot(
         "pal_bre_blue"  = pal_bre_blue(5),
         "material_BRE_blue" = material_BRE_blue
)

scales::show_col(md_bre_blue)
```

## Diverging Color Scale

### Stacked bar chart with `Blue Red 3` color scale:

- Note: This palette was darkened a bit to get the middle color to be a light shade of gray.

```{r}
bre::stackedBarChart(
   df = cat_items_single, fill_color = "div",pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```

## Sequential Color Scales

### Stacked bar chart with `cividis` color scale:

```{r}
bre::stackedBarChart(
   df = cat_items_single, pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```


### Stacked bar chart with white to `BRE` Blue color scale with Colors from Material Design App:

```{r}
bre::stackedBarChart(
   df = cat_items_single, fill_color = material_BRE_blue, pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```

### Stacked bar chart with `BRE` Blue to `BRE` light gray color scale:

```{r}
bre::stackedBarChart(
   df = cat_items_single, fill_color = pal_bre_grey_blue(length(bar_scale_labels)), pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```


### Stacked bar chart with white to `BRE` Blue color scale:

```{r}
bre::stackedBarChart(
   df = cat_items_single, fill_color = pal_bre_blue(length(bar_scale_labels)), pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```


### Stacked bar chart with `Blues 3` color scale:

```{r}
# `Blues 3` from `colorspace` package
bre::stackedBarChart(
   df = cat_items_single, fill_color = colorspace::sequential_hcl(length(bar_scale_labels), "Blues 3", rev = TRUE), pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```

### Stacked bar chart with `Mako` color scale:

```{r}
# `Blues 3` from `colorspace` package
bre::stackedBarChart(
   df = cat_items_single, fill_color = viridis::viridis(n = length(bar_scale_labels), direction = -1, option = "mako"), pre_post = FALSE, scale_labels = bar_scale_labels,
   question_labels = NULL, percent_label = TRUE, width = NULL, overall_n = F
)
```


## Qualitative/Categorical Color Scales

### Arrow chart with `viridis` color scale:

```{r}
bre::arrowChart(df = items, scale_labels = levels_min_ext, arrow_colors = viridisLite::viridis(n = length(levels_min_ext)),
    overall_n = FALSE, question_labels = NULL, question_order = FALSE)
```

### Arrow chart with `Okabe-Ito` color scale:

```{r}
bre::arrowChart(df = items, scale_labels = levels_min_ext, arrow_colors = grDevices::palette.colors(n = length(levels_min_ext), palette = "Okabe-Ito"),
    overall_n = FALSE, question_labels = NULL, question_order = FALSE)
```

### Arrow chart with `turbo` color scale:

```{r}
bre::arrowChart(df = items, scale_labels = levels_min_ext, arrow_colors = viridisLite::turbo(n = length(levels_min_ext)),
    overall_n = FALSE, question_labels = NULL, question_order = FALSE)
```


### Horizontal Bar Chart with `viridis` color scale:

```{r}
# New fake data for testing horzBarChart() colors:
data <- dplyr::tibble(
  role = c(
    "Faculty", "Postdoc", "Undergraduate student", "Graduate student",
    "Graduate student", "Postdoc", "Postdoc", "Faculty",
    "Faculty", "Graduate student", "Graduate student", "Postdoc",
    "Faculty", "Faculty", "Faculty", "Faculty", "Faculty", "Graduate student",
    "Undergraduate student", "Undergraduate student"
  )
)
# set up data for chart:
role_summ <- data %>%
  dplyr::select(role) %>%
  bre::dataSumm()

# horzBarChart() with viridis colors
bre::horzBarChart(df = role_summ, scale_colors = viridisLite::viridis(n = length(unique(data[["role"]]))), width = 0.6)

```

### Horizontal Bar Chart with `Okabe-Ito` color scale:

```{r}
# horzBarChart() with Okabe-Ito colors
bre::horzBarChart(df = role_summ, scale_colors = grDevices::palette.colors(n = length(unique(data[["role"]])), palette = "Okabe-Ito"), width = 0.6)

```
